
\documentclass[a4paper]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.00.0.2552}
%TCIDATA{<META NAME="SaveForMode" CONTENT="1">}
%TCIDATA{Created=Wednesday, September 07, 2005 13:30:08}
%TCIDATA{LastRevised=Sunday, September 11, 2005 10:24:25}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{<META NAME="DocumentShell" CONTENT="Standard LaTeX\Blank - Standard LaTeX Article">}
%TCIDATA{Language=American English}
%TCIDATA{CSTFile=40 LaTeX article.cst}

\setlength{\textheight}{8in}
\setlength{\textwidth}{6in}
\setlength{\oddsidemargin}{0mm}
\setlength{\evensidemargin}{0mm}
\setlength{\marginparwidth}{0mm}
\setlength{\marginparsep}{0mm}
\newtheorem{theorem}{Theorem}
\newtheorem{acknowledgement}[theorem]{Acknowledgement}
\newtheorem{algorithm}[theorem]{Algorithm}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{case}[theorem]{Case}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conclusion}[theorem]{Conclusion}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{criterion}[theorem]{Criterion}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{solution}[theorem]{Solution}
\newtheorem{summary}[theorem]{Summary}
\newenvironment{proof}[1][Proof]{\noindent\textbf{#1.} }{\ \rule{0.5em}{0.5em}}
\input{tcilatex}

\begin{document}

\title{OpenMP Implementation of Gaussian Elimination}
\date{}
\author{Gabriel Dimitriu}
\maketitle

\section{Introduction}

Let the system be 
\begin{equation}
Ax=b  \label{s1}
\end{equation}%
\qquad \qquad Here $A$ is a dense n x n matrix of coefficients such that $%
A[i][j]=a_{i,j}$, b is an n x 1 vector and x is the desired solution vector.

A system of equations is solved in two stages. First, through a series of
algebraic manipulations, the original system of equations is reduced to an
upper triangular system of the form $Ux=y$, where $U$ is a unit
upper-triangular matrix. In the second stage of solving a system of linear
equations, the upper-triangular system is solved for the variables in
reverse order from $x_{n-1}$ to $x_{0}$ by a procedure know as
back-substitution.

\begin{algorithm}
gaussian\_elimination
\end{algorithm}

\begin{enumerate}
\item procedure gaussian\_elimination(A,b,y)

\item begin

\item \qquad for k:=0 to n-1 do

\item \qquad \qquad for j:=k+1 to n-1 do

\item \qquad \qquad \qquad A[k][j]:=A[k][j]/A[k][k]; //division step

\item \qquad \qquad y[k]:=b[k]/A[k][k];

\item \qquad \qquad A[k][k]:=1;

\item \qquad \qquad for i:=k+1 to n-1 do

\item \qquad \qquad \qquad for j:=k+1 to n-1 do

\item \qquad \qquad \qquad \qquad A[i][j]:=A[i][j]-A[i]k]xA[k][j];
//elimination step

\item \qquad \qquad \qquad b[i]:=b[i]-A[i][k]xy[k];

\item \qquad \qquad \qquad A[i][k]:=0;

\item \qquad \qquad endfor;

\item \qquad endfor;

\item end gaussian\_elimination;
\end{enumerate}

\section{\protect\bigskip Pipeline Communication and Computation}

\bigskip The matrix is scattered to all processor so two consecutive row are
in two consecutive processors.

During the $k^{th}$ iteration processor $P_{k}$ broadcast part of the $%
k^{th} $ row of the matrix to processors $P_{k+1},...,P_{p-1}$. Assume that
the processors $P_{0}...P_{p-1}$ are connected in a linear array, and $%
P_{k+1}$ is the first processor to receive the $k^{th}$ row from processor $%
P_{k}$. Then the processor $P_{k+1}$ must forward this data to $P_{k+2}$.
However, after forwarding the $k^{th}$ row to $P_{k+2}$, processor $P_{k+1}$
need not wait to perform the elimination step until all the processors up to 
$P_{p-1}$ have received the $k^{th}\,$\ row. Similarly, $P_{k+2}$ can start
its computation as soon as is has forwarded the $k^{th}$ row to $P_{k+3}$,
and so on. Meanwhile, after completing the computation for the $k^{th}$
iteration, $P_{k+1}$ can perform the division step, and start the broadcast
of the $(k+1)^{th}$ row by sending it to $P_{k+2}$. In this case of shared
memory the receive, send and broadcast from MPI\ are replaced by OpenMP lock
procedures but without pipeline communication.

This algorithm is made for comparing with MPI implementation.

\section{Results}

I have compiled the Gaussian Elimination without load balancing and pivoting
program with two OpenMP compilers: Omni 1.6 and Intel C Compiler 8.0 for
LINUX and the serial with gcc and Intel C Compiler 8.0 for LINUX both with
maximum optimization "-O3" and for Intel C Compiler I've put also
"-mcpu=pentiumpro -tpp6" for maximum optimization.

The executable were run on a dual pentium II at 500MHz with 256MB RAM and
with LINUX\ Fedora Core 1.

The following results were made for a average or 10 runs for serial and
parallel programs and with blue is plotted the results from Omni and with
red the results from ICC.

\begin{center}
\FRAME{itbpF}{4.5994in}{3.4562in}{0in}{}{}{Figure}{\special{language
"Scientific Word";type "GRAPHIC";maintain-aspect-ratio TRUE;display
"USEDEF";valid_file "T";width 4.5994in;height 3.4562in;depth
0in;original-width 5.3756in;original-height 4.0315in;cropleft "0";croptop
"1";cropright "1";cropbottom "0";tempfilename
'full_openmp_ge_2500.bmp';tempfile-properties "XNPR";}}
\end{center}

\end{document}
